The Lipid Interactome Repository

### 2025-01-16 IndividualStudies/DH_2017.qmd
EOD push

---
title: "Höglinger et al. 2017"
subtitle: "Trifunctional lipid probes for comprehensive studies of single lipid species in living cells"
format:
  html:
    embed-resources: false
standalone: false
---

@@

Correspondence: <schulcar@ohsu.edu>{target="_blank"}

@@

>>>>

---
title: "Höglinger et al. 2017"
subtitle: "Trifunctional lipid probes for comprehensive studies of single lipid species in living cells"
format:
  html:
    embed-resources: false
standalone: false
---
<!-- R setup, reading data tables, wrangling data -->
```{r setup, echo=FALSE, include=FALSE}

## intitialize libraries
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(grid)
library(readr)
library(rmarkdown)
library(emojifont)
library(gt)
library(crosstalk)
library(DT)
library(stringr)
library(pheatmap)

DH_2017_data <- read.csv("DataTables/DH_PNAS_2017_download.csv") |>
  mutate(gene_name = str_extract(Description, "GN=([^\\s]+)") |> str_remove("GN=")) |>
  mutate(Species = str_extract(Description, "OS=([^=]+?)\\sGN=") |> str_remove("OS=") |> str_remove("GN=")) |>
  filter(Species == "Homo sapiens ") |>
  mutate(PSM_Sph = (PSM_Sph1 + PSM_Sph2) / 2,
         PSM_FA = (PSM_FA1 + PSM_FA2) / 2,
         PSM_DAG = (PSM_DAG1 + PSM_DAG2) /2) |>
  select(gene_name, PSM_Sph, PSM_FA, PSM_DAG) |>

  glimpse()

```

@@

@@

```{r, echo=FALSE, warning=FALSE, fig.height=15}

DH_2017_data_long <- read.csv("DataTables/DH_PNAS_2017_long.csv")

DH_2017_data_long$name <- factor(pull(DH_2017_data_long, name), ordered = TRUE, levels = c("Sph", "FA", "DAG"))

m <- list(l=50, r = 50, b= 100, pad = 4)

heatmap <- DH_2017_data_long |>
  ggplot(aes(y = id, x = name)) +
  theme(panel.border = element_rect(color = NA, fill = "grey25")) +
  geom_tile(aes(fill = value, text = paste0("Gene name: ", gene_name, '\n', "Lipid Probe:", name, "\n", "PSM: ", value))) +
  scale_fill_gradientn(colors = c("grey95", "purple", "orange"),
  values = scales::rescale(c(0, 10, 25))) +
  ylab("Gene") +
  xlab("Lipid Probe") +
  theme_minimal() +
  theme(text = element_text(family = "serif")) +

  theme(legend.title = element_text(size = 12, face = "bold")) +
  theme(axis.title = element_text(size = 12, face = "bold")) +
  theme(legend.text = element_text(size = 12))

ggplotly(heatmap,
        tooltip = c("text", dynamicTicks = TRUE),
        hovermode = "closest",
        autosize = T,
        margin = m) |>
  layout(showlegend = FALSE)

```

### 2025-01-16 IndividualStudies/RM_2021.qmd
EOD push

---
title: "Farley et al. 2024 B"
subtitle: "Trifunctional fatty acid derivatives: the impact of diazirine placement"
format: 
  html:
    embed-resources: false
standalone: false
---

@@

#### Authors
Scotland E. Farley, Ryu Hashimoto, Judah Evangelista, Frank Stein, Per Haberkant, Kazuya Kikuchi, Fikadu G. Tafesse, Carsten Schultz*

Correspondence: <schulcar@ohsu.edu>{target="_blank"}

#### Journal
Chemical Communications

<https://doi.org/10.1021/acschembio.3c00554>{target="_blank"}

@@

Functionalized lipid probes are a critical new tool to interrogate the function of individual lipid species, but the structural parameters that constrain their utility have not been thoroughly described. Functionalized lipid probes are a critical new tool to interrogate the function of individual lipid species, but the structural parameters that constrain their utility have not been thoroughly described. Here, we synthesize three palmitic acid derivatives with a diazirine at different positions on the acyl chain and examine their metabolism, subcellular localization, and protein interactions. We demonstrate that while they produce very similar metabolites and subcellular distributions, probes with the diazirine at either end pulldown distinct subsets of proteins after photo-crosslinking. This highlights the importance of thoughtful diazirine placement when developing probes based on biological molecules.

@@

@@

[Huh7](https://huh7.com){target="_blank"}

@@

16-channel Tandem Mass Tagging (TMT16)

@@

#### Ranked-order Log2 Fold-change plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Ranked-order plots depict the relative enrichment of each protein versus control (y-axis) from lowest to highest. The left panel depicts the results of a pulldown using trifunctionalized phosphatidic acid; the right panel depicts the results of the pulldown using trifunctionalized phosphatidylethanolamine. Black proteins are unenriched or depleted in the presence of probe, Purple proteins are enriched to the probe but fall below a FDR or p-value threshold of significance, and Orange proteins surpass the threshold of significance (need details of threshold)."

RankedOrderPlotStandard(SF_2024_ranked)
```
:::

#### Volcano plots

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Volcano plots depict the relative enrichment of each protein versus control (x-axis) and the statistical significance of each protein following a Student's t-test (y-axis, -log10 transformed). The left panel depicts the results of a pulldown using trifunctionalized phosphatidic acid; the right panel depicts the results of the pulldown using trifunctionalized phosphatidylethanolamine. Black proteins are unenriched or depleted in the presence of probe, Purple proteins are enriched to the probe but fall below a FDR or p-value threshold of significance, and Orange proteins surpass the threshold of significance (need details of threshold)."

VolcanoPlotStandardized(SF_2024_data)

```
:::

## Enriched Proteins

Check the boxes below to filter the dataset by which lipid probe was used in the pulldown and by significance thresholds.

:::{.border}
```{r, echo=FALSE, warning=FALSE}

# Here is the code to make the filterable lists shown in the crosstalk table below

## Creates a reactive dataframe
SF_2024_data_reactive <- crosstalk::SharedData$new(SF_2024_data)

htmltools::tags$div(
  class = "filter-container",
  filter_checkbox(id = "LipidProbe",
                  label = "Lipid probe in pulldown:", 
                  sharedData = SF_2024_data_reactive,
                  group = ~LipidProbe),
  filter_checkbox(id = "hit_annotation",
                  label = "Enrichment criteria:", 
                  sharedData = SF_2024_data_reactive,
                  group = ~hit_annotation)
)

## Displays the reactive dataframe
datatable(SF_2024_data_reactive, options = list(pageLength = 20))

```
:::

@@

{{< downloadthis DataTables/SF_Chem_Comm_2024_download.csv dname="SF_Chem_Comm_2024">}}

>>>>

---
title: "Muller et al. 2021"
subtitle: "Synthesis and Cellular Labeling of Multifunctional Phosphatidylinositol *Bis*- and *Tris*phosphate Derivatives"
format: 
  html:
    embed-resources: false
standalone: false
---

@@

#### Authors
Rainer Muller, Ana Kojic, Mevlut Citir, Carsten Schultz

#### Journal
Angewandte Chemie

<https://doi.org/10.1002/anie.202103599>{target="_blank"}

@@

We synthesized the first multifunctionalized phosphoinositide polyphosphate derivatives featuring a photo-removable protecting group (“cage”), a photo-crosslinkable diazirine group, and a terminal alkyne group useful for click chemistry. We demonstrate that the lipid derivatives readily enter cells. After photo-crosslinking, cell fixation and fluorescent tagging via click chemistry, we determined the intracellular location of the lipid derivatives before and after uncaging of the lipids. We find that there is rapid trafficking of PI(3,4)P2 and PI(3,4,5)P3 derivatives to the plasma membrane, opening the intriguing possibility that there is active transport of these lipids involved. We employed the photo-crosslinking and click chemistry functions to analyze the proteome of PI(3,4,5)P3-binding proteins. From the latter, we validated by RNAi that the putative lipid binding proteins ATP11A and MPP6 are involved in the transport of PI(3,4,5)P3 to the plasma membrane.

@@

[PI(3,4,5)P~3~](/LipidProbe/PhosphatidylinositolTrisphosphate.qmd)

@@

[HeLa](https://www.atcc.org/products/ccl-2){target="_blank"}

@@

Peptide spectral counting

@@

#### Heat map

#### Venn Diagram

@@

{{< downloadthis DataTables/RM_Angewandte_2021_download.csv dname="RM_Angewandte_2021">}}

### 2025-01-16 LipidProbe/1-10_FattyAcid.qmd
Testing embedded chem structures
EOD push

---
title: "Trifunctional Sphingosine"
---

@@

## Subcellular localization

@@

>>>>

---
title: "Trifunctional 1-10 Fatty Acid"
---

<!-- R setup, reading dataframes and minor data wrangling -->
```{r setup, echo=FALSE, include=FALSE}
## intitialize libraries
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(grid)
library(readr)
library(rmarkdown)
library(emojifont)
library(gt)
library(crosstalk)
library(DT)
library(htmltools)

## Opening the source file for the ggplot custom theme
source("ggplot_styles.R")

# Opening the datatable for this page
SF_2024_data <- readr::read_csv("DataSets/1-10_FA_Huh7_SF_2024.csv") |>
	filter(LipidProbe == "1-10 FA")

# Organizing the hit annotations as factors
SF_2024_data$hit_annotation <- factor(pull(SF_2024_data, hit_annotation), levels = c("enriched hit", "hit", "enriched candidate", "candidate", "no hit", "NA"), ordered = TRUE)

```

@@

@@

::: {.panel-tabset}

## Huh7 cells

```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Data as reported in [Farley et al., 2024(B)](/IndividualStudies/SF_2024_B.qmd)"
VolcanoPlotStandardized(SF_2024_data)
```

:::

## Enriched hits

```{r, echo=FALSE}

```

### 2025-01-16 LipidProbe/8-3_FattyAcid.qmd
Testing embedded chem structures
EOD push

---
title: "Trifunctional 1-10 Fatty Acid"
---

>>>>

---
title: "Trifunctional 8-3 Fatty Acid"
---

<!-- R setup, reading dataframes and minor data wrangling -->
```{r setup, echo=FALSE, include=FALSE}
## intitialize libraries
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(grid)
library(readr)
library(rmarkdown)
library(emojifont)
library(gt)
library(crosstalk)
library(DT)
library(htmltools)

## Opening the source file for the ggplot custom theme
source("ggplot_styles.R")

# Opening the datatable for this page
SF_2024_data <- readr::read_csv("DataSets/8-3FA_Huh7_SF_2024.csv") |>
	filter(LipidProbe == "8-3 FA")

# Organizing the hit annotations as factors
SF_2024_data$hit_annotation <- factor(pull(SF_2024_data, hit_annotation), levels = c("enriched hit", "hit", "enriched candidate", "candidate", "no hit", "NA"), ordered = TRUE)

```

### 2025-01-16 index.qmd
EOD push

@@

@@

    <br>
    <a href="LipidProbe/Diacylglycerol.qmd" style="text-decoration: none;">
        <button style="background-color: #3277a8; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            Diacylglycerol
        </button>
    </a>
		<br>
    <a href="LipidProbe/Sphinganine.qmd" style="text-decoration: none;">
        <button style="background-color: #2ecc71; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            Sphinganine
        </button>
    </a>
		<a href="LipidProbe/Sphingosine.qmd" style="text-decoration: none;">
        <button style="background-color: #eba434; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            Sphingosine
        </button>
    </a>

	<a href="LipidProbe/8-3_FattyAcid.qmd" style="text-decoration: none;">
        <button style="background-color: #ba34eb; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            8-3 Fatty Acid
        </button>
    </a>

>>>>

## Site Overview

@@

## Lipid Interactomes

@@

    <br>
    <a href="LipidProbe/Diacylglycerol.qmd" style="text-decoration: none;">
        <button style="background-color: #3277a8; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            Diacylglycerol
        </button>
    </a>

    <br>
    <a href="LipidProbe/Sphinganine.qmd" style="text-decoration: none;">
        <button style="background-color: #2ecc71; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            Sphinganine
        </button>
    </a>
    <a href="LipidProbe/Sphingosine.qmd" style="text-decoration: none;">
        <button style="background-color: #eba434; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            Sphingosine
        </button>
    </a>

    <br>
	<a href="LipidProbe/8-3_FattyAcid.qmd" style="text-decoration: none;">
        <button style="background-color: #ba34eb; color: white; font-size: 20px; padding: 15px 30px; border: none; border-radius: 8px; margin: 10px; cursor: pointer;">
            8-3 Fatty Acid
        </button>
    </a>
